Lipschitz constrained networks have gathered considerable attention in the deep learning community, with usages ranging from Wasserstein distance estimation to the training of certifiably robust classifiers. However they remain commonly considered as less accurate, and their properties in learning are still not fully understood. In this paper we clarify the matter: when it comes to classification 1-Lipschitz neural networks enjoy several advantages over their unconstrained counterpart. First, we show that these networks are as accurate as classical ones, and can fit arbitrarily difficult boundaries. Then, relying on a robustness metric that reflects operational needs we characterize the most robust classifier: the WGAN discriminator. Next, we show that 1-Lipschitz neural networks generalize well under milder assumptions. Finally, we show that hyper-parameters of the loss are crucial for controlling the accuracy-robustness trade-off. We conclude that they exhibit appealing properties to pave the way toward provably accurate, and provably robust neural networks.    
% \todo{a mon gout c'est risqué de parler de ça des l'abstract: brouille le message qu'on traite uniquement de classification.}
% The need for certifiable robustness to adversarial attacks have drawn a considerable attention on Lipschitz constrained networks. %in the deep learning community.
% Despite that, properties of such networks are still not fully understood, and they remain commonly considered as less expressive than their unconstrained counterpart.
% In this paper we challenge this misconception, and we show instead that 1-Lipschitz neural networks enjoy several appealing properties over classical networks in classification.
% First, we show they can fit arbitrarily difficult boundaries, making them as accurate as classical ones. 
% Then, we characterize the maximally robust classifier, and a robustness metric which reflect operational needs.
% We show common losses exhibit (usually omitted) hyper-parameters that control the trade-off between accuracy and robustness.
