\chapter{Background on Lipschitz Neural Networks}
\label{ch:background}

\section{Lipschitz Continuity in Deep Learning}

    \subsection{Lipschitz Neural Networks are expressive for classification}

    \subsection{Robustness}

    paper "achieving robustness"

    \subsection{Explainability}

    paper "explainable 1 lip" + paper "beyond adversarial robustness"

    \subsection{Differential privacy}

    paper "dp-sgd without clipping"

\section{Building Lipschitz Neural Networks}

    \subsection{On the evaluation of Lipschitz constant}

    \subsection{Global strategy}

    \subsection{Product bound: both a blessing and a curse?}


\chapter{From Lipschitzness to Orthogonality Constraints}
\label{ch:background}

\section{Why enforcing orthogonality on weights?}

    Orthogonal layers have become fundamental components in various deep learning architectures due to their unique mathematical properties, which offer benefits across multiple applications. 
    For instance, robustness against adversarial attacks can be achieved by managing a model's Lipschitz constant~\cite{szegedy_intriguing_2013} -- with 1-Lipschitz networks being a prime candidate: such networks allowing the computation of robustness certificates. 
    Initially, researchers experimented with regularization techniques~\cite{cisse2017parseval}; however, constrained networks, especially those employing orthogonal layers, soon became central, as they provided the advantage of tighter certification bounds: The overall Lipschitz constant of a sequence of layers is typically estimated as the product of the individual layer constants; however, this bound is often loose, and computing the exact constant is NP-hard \cite{virmaux2018lipschitz}. \cite{anil_sorting_2019} 
    shows that combining MaxMin activation with orthogonal layers makes the aforementioned bound tight.
    Beyond robustness, orthogonal layers also play a key role in enhancing performance in normalizing flows. Normalizing flows are generative models that transform simple distributions into complex ones via invertible mappings \cite{dinh2016density, rezende2015variational}.  Orthogonal convolutions enable these transformations with a computable Jacobian determinant, thus improving training efficiency \cite{kingma2018glow} and forming the basis for invertible residual networks \cite{behrmann2019invertible}.
    Additionally, orthogonal layers stabilize training deep and recurrent neural networks (RNNs) by preserving gradient norms through time, essential in capturing long-term dependencies in time-series, such as language and speech tasks \cite{kiani_projunn_2022,qi_deep_isometric_2020, bansal2018gainorthogonalityregularizationstraining}. Lastly, in Wasserstein GANs (WGANs) \cite{arjovsky2017wasserstein} orthogonality in both the discriminator and generator~\cite{muller2019orthogonal} supports stability and expressivity without requiring weight clipping or gradient penalties~\cite{gulrajani2017improved}, making it essential for large-scale GAN training~\cite{brock2018large, miyato2018spectral}.


    
    \paragraph{Provable Robustness with 1-Lipschitz Networks.}
    Ensuring robustness against adversarial attacks is a critical challenge. Early work in this field \cite{szegedy_intriguing_2013} identified a link between a network's adversarial robustness and its Lipschitz constant, which led to the development of networks with a Lipschitz constant of one (1-Lipschitz networks). Initially, regularization techniques were used \cite{cisse2017parseval}, but interest in constrained networks quickly grew. Orthogonal layers, in particular, have an inherent Lipschitz constant of one, as they preserve the input norm through each transformation. This property is instrumental in achieving provable robustness by allowing for certified bounds on the network's output perturbations in response to adversarial inputs \cite{anil_sorting_2019}. By controlling the network's sensitivity to input changes, orthogonal layers play a crucial role in building models resilient to adversarial manipulations.
    Besides robustness, Lipschitz-constrained networks have diverse applications: they are closely linked to WGANs and optimal transport \cite{arjovsky2017wasserstein, bethune2024deep}, enable scalable differential privacy \cite{bethunedp}, allow conformal prediction in adversarial setting \cite{massena2025efficient}, and prevent singularities in diffusion models \cite{yang2023lipschitz}. Additionally, they guarantee the existence and uniqueness of solutions in classification \cite{bethune_pay_2022} and flow-matching networks \cite{perko2013differential, coddington1956theory}.
    
    \paragraph{Enhancing Performance in Normalizing Flows.}
    Normalizing flows are a class of generative models that transform simple probability distributions into complex ones through a series of invertible and differentiable mappings. Although they have different objectives, this domain intersects with the field of provable adversarial robustness. For instance, \cite{dinh2016density, rezende2015variational} employs a channel masking scheme, which was later used to emulate striding in Lipschitz layers designed for adversarial robustness. Separately, Lipschitz layers can be applied to build invertible residual networks \cite{behrmann2019invertible}. In both fields, orthogonal convolutions are essential, as they facilitate the construction of invertible transformations with tractable Jacobian determinants. The use of orthogonal layers ensures that the Jacobian determinant is constant or easily computable, simplifying likelihood estimation during training \cite{kingma2018glow}. This property enables efficient and stable training of normalizing flow models, leading to improved performance in density estimation and generative tasks.
    
    \paragraph{Stabilizing Training in Deep and Recurrent Neural Networks.}
    Training recurrent neural networks (RNNs) involves propagating gradients through time, which can lead to vanishing or exploding gradients due to the multiplicative nature of sequential weight applications. Orthogonal weight matrices in RNNs help preserve the gradient norm across time steps, thus preventing degradation of the learning signal \cite{kiani_projunn_2022}. By constraining recurrent weights to be orthogonal, the network maintains a consistent flow of information, enabling it to capture long-term dependencies more effectively. This stabilization is essential for tasks requiring understanding long-range dependencies in time series, such as language modeling and speech recognition. These approaches also facilitate the training of very deep networks \cite{qi_deep_isometric_2020} with improved generalization properties \cite{bansal2018gainorthogonalityregularizationstraining}.
    
    \paragraph{Improving Stability in Wasserstein Generative Adversarial Networks.}
    Generative Adversarial Networks (GANs) are powerful models for generating realistic data, but they often suffer from training instability. Wasserstein GANs (WGANs) \cite{arjovsky2017wasserstein} address this issue by optimizing the Wasserstein distance between the real and generated data distributions. A key requirement for WGANs is that the discriminator (or critic) function must be Lipschitz continuous. Orthogonal layers naturally satisfy this Lipschitz condition, eliminating the need for techniques like weight clipping or gradient penalties \cite{gulrajani2017improved}, which can adversely affect training dynamics. By incorporating orthogonal convolutions into the discriminator, WGANs achieve more stable training \cite{miyato2018spectral, muller2019orthogonal} and produce higher-quality generative results. Orthogonality has been integral to the successful scaling of GAN training \cite{brock2018large}. %\textbf{https://github.com/ezqlwyb/GAN}

    \paragraph{Extending the applicability of Proximal Neural Networks.}
    Proximal Neural Networks (PNNs) offer a principled approach to designing and training deep architectures by drawing inspiration from optimization theory—specifically, proximal algorithms and $\alpha$-averaged operators \cite{hasannasab2020parseval, combettes2020lipschitz}. A central insight in this framework is the emergence and enforcement of orthogonality in learned operators, which can improve network stability. \cite{hertrich2021convolutional} proposes an extension to convolutional operators but notes that, for limited kernel sizes, optimizing on the manifold is challenging and instead relies on penalization, similar to \cite{wang_orthogonal_2020}.

\section{Convolutions and orthogonality}

    Despite these benefits, extending orthogonality to convolutional layers remains challenging: on one side, 2D matrices and convolutions are both linear operators, making orthogonality well defined. On the other side, a convolution corresponds to a large, structured linear operator with Toeplitz (or doubly circulant) structure. Orthogonalizing such matrices without breaking their convolutional structure (sparsity, weight sharing, locality) is delicate, and in some regimes provably infeasible~\cite{achour2023existencestabilityscalabilityorthogonal}. Beyond practical considerations, controlling these structured operators has theoretical importance: it affects generalization~\cite{bethune_pay_2022} and dictates when an orthogonal convolution can even exist~\cite{achour2023existencestabilityscalabilityorthogonal}.

    Early approaches relied on \emph{regularization}: \cite{wang_orthogonal_2020} penalizes deviations from orthogonality of the doubly block circulant matrix associated with the kernel, while \cite{qi_deep_isometric_2020} adds an isometry-promoting term during training. These methods are simple and architecture-agnostic, but they offer no exact guarantees: the resulting layers are only approximately 1-Lipschitz, the penalty interferes with the task loss, and the strength of the regularizer must be tuned for each architecture. These limitations motivated a line of work on \emph{explicit} parameterizations that guarantee orthogonality by construction, which we now review and group by the strategy they adopt.

    \subsection{Explicit construction in the spatial domain}

        \paragraph{BCOP (Block Convolution Orthogonal Parameterization).}
        Building on the 1D orthogonal convolution construction of \cite{xiao_dynamical_2018}, BCOP~\cite{li_preventing_2019} parameterizes a $k \times k$ orthogonal kernel as the composition (via block convolution) of $2k-1$ elementary $1 \times 2$ or $2 \times 1$ convolutions, each built from a pair of symmetric projectors obtained from orthogonalized rectangular matrices. The composition yields a kernel that is provably orthogonal in the spatial domain and admits an arbitrary kernel size at no extra runtime cost (the composed kernel is materialized once). Its main strengths are an explicit kernel representation (compatible with standard convolution operators and therefore cheap at inference) and a tight effective receptive field equal to $k$. Its weaknesses are well documented: the parameterization is provably incomplete (it does not cover the full manifold of orthogonal kernels of size $k$~\cite{li_preventing_2019}), it requires $c_i \geq c_o$ (or padding), and the original implementation supports neither native striding, nor transposed, dilated, or grouped convolutions.

        \paragraph{SC-Fac (Scaling-up via a paraunitary framework).}
        SC-Fac~\cite{su_scaling-up_2021} reinterprets BCOP through the lens of paraunitary matrices from signal processing and shows that orthogonal convolutions can be factorized into separable $1\times k$ then $k \times 1$ orthogonal blocks. This factorization yields a richer family of kernels than BCOP and supports stride, dilation and groups, while keeping a spatial-domain construction. The price is twofold: the resulting effective kernel is separable -- which restricts expressivity compared to a full $k\times k$ kernel -- and at the time of writing no public, well-maintained implementation exists, which has hindered its empirical adoption.

        \paragraph{ECO (Explicitly Constructed Orthogonal convolutions).}
        ECO~\cite{yu_constructing_2021} parameterizes the kernel through a small set of $c \times c$ skew-symmetric matrices, applies a matrix exponential channel-wise in the Fourier basis, and reconstructs the spatial kernel via an inverse 2D DFT. To control the number of non-zero parameters, the kernel is materialized as a \emph{dilated} convolution with dilation $n/k$, so that only $k\times k$ entries are non-zero. This yields a strict orthogonal layer with controlled parameter count and a true local receptive field. However, the dilation must divide the input resolution $n$, which couples kernel design to image size, and the construction does not natively support striding, transposed or grouped convolutions.

    \subsection{Frequency-domain parameterizations}

        \paragraph{Cayley convolutions.}
        \cite{trockman_orthogonalizing_2021} exploit the fact that a circular multichannel convolution is block-diagonalized by the 2D DFT into $h \times w$ independent $c \times c$ complex matrices. Each block is then made unitary by the Cayley transform $A \mapsto (I - A)(I + A)^{-1}$ applied to a skew-Hermitian parameter, and the inverse FFT yields a real orthogonal convolution. The construction is exact, the per-block Cayley transform is differentiable and well-conditioned, and the parameterization covers the full space of circular orthogonal convolutions. The drawbacks come from operating in the frequency domain: the equivalent kernel has the size of the input image (which inflates the receptive field and the parameter count at large resolutions), the parameterization is tied to circular padding, and channel changes, striding, dilations and grouped convolutions must be emulated through ad-hoc tricks (zero-padding, invertible downsampling) that increase the computational cost.

        \paragraph{LOT (Layer-wise Orthogonal Training).}
        LOT~\cite{xu_lot_2023} follows the same Fourier strategy as Cayley but replaces the Cayley transform with the polar factor $W = (VV^\top)^{-1/2}V$, computed in each frequency block via a differentiable Newton--Schulz iteration. This avoids the matrix inverse required by Cayley and tends to be better conditioned, often translating into improved certified accuracy at equal model size. The inherited limitations are essentially the same: input-sized effective kernel, dependence on the FFT, and no native support for stride, dilation or grouped convolutions. Each forward pass also requires several Newton--Schulz iterations per spatial frequency, which makes LOT one of the most expensive orthogonal layers per call.

        \paragraph{ProjUNN-T.}
        Originally designed for recurrent networks, ProjUNN-T~\cite{kiani_projunn_2022} parameterizes a unitary matrix via a low-rank projection step and was extended to convolutions by applying the same machinery to the per-frequency $c \times c$ blocks obtained by FFT. The low-rank update keeps each optimization step inexpensive even for large channel counts, which makes it attractive when $c$ dominates the cost. In its convolutional form, however, it inherits the frequency-domain restrictions and does not support stride, dilation, groups or changes in the number of channels, which has limited its use in convolutional networks beyond proof-of-concept settings.

        \paragraph{Singular Value Clipping (SVCM).}
        Although somewhat orthogonal to the rest of this list, the singular value clipping of \cite{sedghi_singular_2019} exploits the same FFT block-diagonalization to compute and clip the singular values of a convolution exactly. It yields a 1-Lipschitz (but not orthogonal) projection and produces a kernel that is generically of size $n \times n$, requiring alternating projections onto $k \times k$ kernels to remain compatible with standard convolution. The cost of these alternating projections (around fifty iterations are typically needed) is the main practical obstacle.

    \subsection{Composite and exponential constructions}

        \paragraph{SOC (Skew Orthogonal Convolutions).}
        SOC~\cite{singla_skew_2021} relies on the identity $\exp(J)\,\exp(J)^\top = I$ for any skew-symmetric $J$. They build a skew-symmetric \emph{convolution} via $L = M - \mathrm{conv\_transpose}(M)$ and approximate the convolution exponential $L \star^e x = x + Lx + \tfrac{1}{2}L\star L\star x + \cdots$ by a truncated Taylor series, with a spectral-normalization step on $L$ to control the truncation error. The construction is strictly orthogonal in the limit and has been one of the strongest methods for certified robustness, particularly in deeper networks. Its weaknesses are computational: each forward pass triggers a sequence of convolutions (typically 6 during training, 12 at inference), the effective receptive field grows with the number of Taylor terms (the layer behaves as a convolution of equivalent size $k + n\lceil k/2 \rceil$), and memory consumption is high. Stride, transposition and grouped convolutions are not natively supported.

        \paragraph{Sandwich.}
        \cite{wang_direct_2023} derive a direct parameterization of Lipschitz-bounded networks from an integral quadratic constraint satisfied by a whole \emph{stack} of layers. A Cholesky-type factorization of the resulting linear matrix inequality yields a per-layer structure $W_k = \sqrt{2}A_k^\top \Psi_k\,\sigma\,\sqrt{2}\Psi_k^{-1}B_k$, where $A_k$, $B_k$ are orthogonal (parameterized in Fourier space via the Cayley transform) and $\Psi_k$ is a diagonal rescaling. The resulting block is provably 1-Lipschitz and tends to deliver stronger empirical robustness than naive layer-wise orthogonalization, because the constraint is enforced jointly across consecutive layers. The drawback is that each ``layer'' is in fact a composite of two FFT-based orthogonal convolutions interleaved with the activation, which doubles the per-layer cost and inherits the frequency-domain limitations on stride, dilation, and grouped convolutions.

    \subsection{Relaxations: 1-Lipschitz layers without strict orthogonality}

        A complementary line of work gives up strict orthogonality and instead enforces only the 1-Lipschitz constraint -- often with a side benefit of mitigating vanishing gradients while remaining cheap to evaluate.

        \paragraph{AOL (Almost-Orthogonal Layer).}
        AOL~\cite{prach_almost-orthogonal_2022} rescales each input channel by a quantity computed directly from the kernel itself: $d_c = \big(\sum_{i,j,a,b} |P^{(a,b)} \star P^{(c,b)}|_{i,j}\big)^{-1/2}$. The resulting layer is provably 1-Lipschitz, requires no iterative procedure, and reduces to a single convolution at inference. It is by far the cheapest construction in this family. The downside is that the rescaling is conservative: AOL layers are typically only $\varepsilon$-orthogonal, which hurts expressivity and certified accuracy compared to truly orthogonal alternatives, especially in deeper networks.

        \paragraph{CPL (Convex Potential Layers).}
        Building on dynamical-system analyses of neural networks, CPL~\cite{meunier_dynamical_2022} parameterizes residual blocks of the form $x + W^\top \sigma(Wx + b)$ with an explicit scaling that guarantees the block is 1-Lipschitz. The construction is closed-form, requires no iteration, and is straightforward to plug into convolutional architectures. As with AOL, the price is a loss of orthogonality and a tendency to underperform truly orthogonal methods on certified accuracy at fixed depth.

        \paragraph{SLL (SDP-based Lipschitz Layers).}
        SLL~\cite{araujo_unified_2023} unifies a large class of 1-Lipschitz parameterizations -- including AOL, CPL and spectral normalization -- by deriving them as analytical solutions of a layer-wise semidefinite programming feasibility condition. From this viewpoint, the authors propose a new family of layers with looser, learnable constraints, achieving a better expressivity/cost trade-off than AOL or CPL while remaining fully closed-form. SLL is currently one of the strongest cheap 1-Lipschitz layers, but it shares their fundamental limitation: it is not strictly orthogonal, and the layer is composite (it interleaves linear maps and activations), which makes it hard to compose with modern convolutional features such as stride or transposition.